---
title: "PS3"
author: "Shuai Yan"
date: "2023-10-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*To create reproducible results, set random seed to 1.*
```{r set random seed}
set.seed(19710822)
```

## Part I
```{r model comparison}
myoglobin <- data.frame(s = c(1.1,  1.5,  1.6,  2.3,  3.4,  5.3,  7.5,  8.4, 14.1),
                        v = c(1.49, 1.79, 1.79, 2.11, 2.83, 3.42, 3.79, 3.97, 4.08))

# models
michaelis_menten_model <- nls(v ~ Vmax * s / (Km + s),
                        start = list(Vmax = 4, Km = 2),
                        data = myoglobin)

non_specific_binding_model <- nls(v ~ (Vmax * s / (Km + s)) + Mns * s,
                              start = list(Vmax = 4, Km = 2, Mns = 1),
                              data = myoglobin)

background_signal_model <- nls(v ~ (Vmax * s / (Km + s)) + Vbk,
                              start = list(Vmax = 4, Km = 2, Vbk = 1),
                              data = myoglobin)

ns_bk_model <- nls(v ~ (Vmax * s / (Km + s)) + Mns*s + Vbk,
                     start = list(Vmax = 4, Km = 2, Mns = 0.5, Vbk = 0.5),
                     data = myoglobin)

# AIC correction
corr_AIC <- function(AIC_value, n, P) {
  return(AIC_value + 2 * (P + 1) * (P + 2) / (n - P))
}

# corrected AIC
n <- length(myoglobin$s)
P_mm <- length(coefficients(michaelis_menten_model))
P_ns <- length(coefficients(non_specific_binding_model))
P_bk <- length(coefficients(background_signal_model))
P_ns_bk <- length(coefficients(ns_bk_model))

AIC_mm <- corr_AIC(AIC(michaelis_menten_model), n, P_mm)
AIC_ns <- corr_AIC(AIC(non_specific_binding_model), n, P_ns)
AIC_bk <- corr_AIC(AIC(background_signal_model), n, P_bk)
AIC_ns_bk <- corr_AIC(AIC(ns_bk_model), n, P_ns_bk)

delta_ns <- AIC_ns - AIC_mm
delta_bk <- AIC_bk - AIC_mm
delta_ns_bk <- AIC_ns_bk - AIC_mm
```

Compared to the Michaelis Menten model, $\Delta AIC$ of the non-specific binding model is: `r delta_ns`; $\Delta AIC$ of the background signal model is: `r delta_bk`; $\Delta AIC$ of the model with both terms is: `r delta_ns_bk`. None of them is less then -6, thus there is not enough evidence to choose either of the more complex models.

## Part II
```{r multiple T test}
control <- data.frame(    t = rnorm(10, mean = 7,  sd = 0.6),
                      group = factor("ctrl"))
drug_a  <- data.frame(    t = rnorm( 8, mean = 9,  sd = 0.6),
                      group = factor("drgA"))
drug_b  <- data.frame(    t = rnorm( 9, mean = 7,  sd = 0.6),
                      group = factor("drgB"))
drug_c  <- data.frame(    t = rnorm( 7, mean = 7,  sd = 0.6),
                      group = factor("drgC"))
drug_d  <- data.frame(    t = rnorm( 8, mean = 11, sd = 0.6),
                      group = factor("drgD"))
d <- rbind(control, drug_a, drug_b, drug_c, drug_d)

a_test <- t.test(x=drug_a$t, y=control$t, conf.level=0.05)
b_test <- t.test(x=drug_b$t, y=control$t, conf.level=0.05)
c_test <- t.test(x=drug_c$t, y=control$t, conf.level=0.05)
d_test <- t.test(x=drug_d$t, y=control$t, conf.level=0.05)
```
```{r A}
print(a_test)
```

```{r B}
print(b_test)
```

```{r C}
print(c_test)
```

```{r D}
print(d_test)
```
Thus drug A, drug B, and drug D are significant.

```{r multiple T test correction}
p_raw <- c(a_test$p.value,
           b_test$p.value,
           c_test$p.value,
           d_test$p.value)
print(which(p.adjust(p_raw, method="BH") < 0.05))
```
Thus, our result is not changed after Benjamini Hochberg correction.

```{r ANOVA}
anova_result <- aov(t ~ group, data=d)
tHSD <- TukeyHSD(anova_result)
print(tHSD)

# plot
par(mar=c(5,8,2,1)) ; plot(tHSD, las=1) # horizontal x-axis labels
abline(v = c(-2.5, 2.5), col="red")
```
Only drug A and drug D are significant based on ANOVA and Tukey’s Honest Significant Differences post-testing procedure.

Based on our calculation, I prefer omnibus test followed by post-testing procedure. The reason is that this test provides more power over multiple T tests, which is mostly persued except for screening. Only if the probem I am facing is clearly a screening problem, in which I want more positive data regardless of more false positive, I may use multiple T tests instead. In addition, there is a higher risk of forgetting correction after multiple T tests, which is not a problem with my preference.
